---
title: "BasicDataAnalysis"
author: "Matt Cliff"
date: "January 29, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = NA)
require(tidyverse)
require(xlsx)
require(dplyr)
require(ggplot2)
```

## Basic Data Analysis

This report focused on reading several data sets, performing exploratory data analysis and saving a tidy data set to use for modeling.

List of [https://vincentarelbundock.github.io/Rdatasets/datasets.html](Data Sources)

We want to find something with at least 10 features, and 5000 observations
[https://archive.ics.uci.edu/ml/datasets.html?format=&task=cla&att=&area=&numAtt=10to100&numIns=&type=&sort=instDown&view=table]

* [https://archive.ics.uci.edu/ml/datasets/Buzz+in+social+media+](Buzz in Social Media)

* [Sensorless drive](https://archive.ics.uci.edu/ml/datasets/Dataset+for+Sensorless+Drive+Diagnosis) - 49 attributes, 11 classes, 58000 records 
* [parkinsons](https://archive.ics.uci.edu/ml/datasets/Parkinsons) binary classifier with 23 predictors; 197 instances
* [card payment](https://archive.ics.uci.edu/ml/datasets/default+of+credit+card+clients) binary classifier with 23 predictors


### Credit Payment Dataset
This comes from [card payment](https://archive.ics.uci.edu/ml/datasets/default+of+credit+card+clients) binary classifier with 23 predictors

```{r loaddata, echo=FALSE}

ccDataURL <- "https://archive.ics.uci.edu/ml/machine-learning-databases/00350/default%20of%20credit%20card%20clients.xls"
ccDataFile <- "./data/ccDataFile.xls"
dir.create("data", showWarnings=FALSE)
if (!file.exists(ccDataFile)) {
    download.file(url=ccDataURL, destfile=ccDataFile, mode="wb")
}
ccData <- read.xlsx2(file=ccDataFile, sheetIndex=1, startRow=2, stringsAsFactors=FALSE)


```

The basic approach will be to ensure no zero elements
NAndx <- sapply(data, function(x) mean(is.na(x))) > .9
nzv <- nearZeroVar(data)

```{r pulldata, echo=FALSE, warning=FALSE}

fCols <- c("SEX", "EDUCATION", "MARRIAGE", "default.payment.next.month")
nCols <- c("LIMIT_BAL", "AGE", "PAY_0", "PAY_2", "PAY_3", "PAY_4", "PAY_5", "PAY_6",
           "BILL_AMT1", "BILL_AMT2", "BILL_AMT3", "BILL_AMT4", "BILL_AMT5", "BILL_AMT6", 
           "PAY_AMT1", "PAY_AMT2", "PAY_AMT3", "PAY_AMT4", "PAY_AMT5", "PAY_AMT6" )
ccData <- select(ccData, -(ID)) %>% 
    mutate_at(fCols, funs(factor(.))) %>%
    mutate_at(nCols, funs(as.numeric(.)))

# go ahead and write it out for future notebooks
save(ccData, file="data/ccData.RData")
summary(ccData)
```

Some tables
```{r ccTables}
hist(ccData$LIMIT_BAL)

#hist(ccData$SEX)
#hist(ccData$MARRIAGE)
hist(ccData$AGE)

hist(ccData$PAY_0)
```

Here we can see that if the bill amounts are much higher, more likely to default
```{r plot}
#ggplot(ccData, aes(PAY_AMT1, fill=MARRIAGE)) +
#    geom_density(alpha=0.3)
qplot(PAY_AMT1,BILL_AMT1,data=ccData, col=default.payment.next.month)
qplot(PAY_AMT2,BILL_AMT2,data=ccData, col=default.payment.next.month)
qplot(PAY_AMT3,BILL_AMT3,data=ccData, col=default.payment.next.month)

```

Lets look at the table of SEX, Marriage, education to the *default.payment.next.month* indicator.



### Parkinsons

This dataset is composed of a range of biomedical voice measurements from 
31 people, 23 with Parkinson's disease (PD). Each column in the table is a 
particular voice measure, and each row corresponds one of 195 voice 
recording from these individuals ("name" column). The main aim of the data 
is to discriminate healthy people from those with PD, according to "status" 
column which is set to 0 for healthy and 1 for PD.

 There are around six 
recordings per patient, the name of the patient is identified in the first 
column.


(https://archive.ics.uci.edu/ml/datasets/Parkinsons) binary classifier with 23 predictors; 197 instances


Matrix column entries (attributes):
name - ASCII subject name and recording number
MDVP:Fo(Hz) - Average vocal fundamental frequency
MDVP:Fhi(Hz) - Maximum vocal fundamental frequency
MDVP:Flo(Hz) - Minimum vocal fundamental frequency
MDVP:Jitter(%),MDVP:Jitter(Abs),MDVP:RAP,MDVP:PPQ,Jitter:DDP - Several 
measures of variation in fundamental frequency
MDVP:Shimmer,MDVP:Shimmer(dB),Shimmer:APQ3,Shimmer:APQ5,MDVP:APQ,Shimmer:DDA - Several measures of variation in amplitude
NHR,HNR - Two measures of ratio of noise to tonal components in the voice
status - Health status of the subject (one) - Parkinson's, (zero) - healthy
RPDE,D2 - Two nonlinear dynamical complexity measures
DFA - Signal fractal scaling exponent
spread1,spread2,PPE - Three nonlinear measures of fundamental frequency variation 

status - Health status of the subject (one) - Parkinson's, (zero) - healthy
```{r loadPKdata, echo=FALSE}

pkDataURL <- "https://archive.ics.uci.edu/ml/machine-learning-databases/parkinsons/parkinsons.data"
pkDataFile <- "./data/parkinsons.data"
dir.create("data", showWarnings=FALSE)
if (!file.exists(pkDataFile)) {
    download.file(url=pkDataURL, destfile=pkDataFile)
}
pkData <- read.csv(file = pkDataFile)
pkData <- select(pkData, -(name)) %>%
    mutate(status = factor(status))
summary(pkData)
```

Let's take a look at some plots
```{r plotPK}
qplot(MDVP.Fo.Hz., fill=status, alpha=I(.5), data=pkData)
qplot(MDVP.Jitter.Abs., fill=status, alpha=I(.5), geom="density", data=pkData)

qplot(Jitter.DDP, fill=status, alpha=I(.5), geom="density", data=pkData)

qplot(MDVP.Shimmer, fill=status, alpha=I(.5), geom="density", data=pkData)

```

#### References

Max A. Little, Patrick E. McSharry, Eric J. Hunter, Lorraine O. Ramig (2008), 
'Suitability of dysphonia measurements for telemonitoring of Parkinson's disease', 
IEEE Transactions on Biomedical Engineering (to appear).

'Exploiting Nonlinear Recurrence and Fractal Scaling Properties for Voice Disorder Detection', 
Little MA, McSharry PE, Roberts SJ, Costello DAE, Moroz IM. 
BioMedical Engineering OnLine 2007, 6:23 (26 June 2007)




## Save Formatted Data
Finally Save the datasets
```{r writeData}
save(ccData, file="./data/ccData.RData")
save(pkData, file="./data/pkData.RData")

```