---
title: "BasicModelAnalysis"
author: "Matt Cliff"
date: "January 29, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Basic Model Analysis

My goal is to set up a page that I can inject a dataset into
then run numerous models to exhaustion to identify to best one or ones to tune.

To begin with I will need to limit the problem domain based on regression or classification methods, although the underlying mechanics is almost always linear algebra the approaches to these problems can diverge significantly.

### Classification Problem

The Classification (value) problem is defined as one where a particular observation of $n$ predictors will have a single valued dependent (predicted) variable that is from a finite class.  The predictor or indepdenent variables can be either finite valued or real valued (or vector valued).

Our goal is to come up with some model function $f: \mathbb{X}_1 \times \mathbb{X}_2\times \dots \times \mathbb{X}_n\mapsto \mathbb{Y}$ that matches our data and we think can best be used to predict the dependent variable for future observations from the domain space.

Each of the $\mathbb{X_j}$ could be discrete or continuous, we will not have any practicle restriction except that they MUST be valued, any missing data must be dealt with prior to using this package.

We want to find something with at least 10 features, and 5000 observations
[https://archive.ics.uci.edu/ml/datasets.html?format=&task=cla&att=&area=&numAtt=10to100&numIns=&type=&sort=instDown&view=table]

* [https://archive.ics.uci.edu/ml/datasets/Buzz+in+social+media+](Buzz in Social Media)

* [Sensorless drive](https://archive.ics.uci.edu/ml/datasets/Dataset+for+Sensorless+Drive+Diagnosis) - 49 attributes, 11 classes, 58000 records

* [card payment](https://archive.ics.uci.edu/ml/datasets/default+of+credit+card+clients) binary classifier with 23 predictors

* [parkinsons](https://archive.ics.uci.edu/ml/datasets/Parkinsons) binary classifier with 23 predictors; 197 instances

```{r loaddata}
#load("data/ccData.RData")
data("possum", package="DAAG")

```

The basic approach will be to ensure no zero elements
NAndx <- sapply(data, function(x) mean(is.na(x))) > .9
nzv <- nearZeroVar(data)

```{r pulldata}


```

## Data Summary

You can also embed plots, for example:


Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

Q what correlates most to Sex?

## Plots

```{r plot}
#require(ggplot2)
#qplot(possum, x=eye, y=totlngth, col=sex, size=log(age))

```


