---
title: "BasicDataAnalysis"
author: "Matt Cliff"
date: "January 29, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = NA)
require(tidyverse)
require(xlsx)
require(dplyr)
require(ggplot2)
```

## Basic Data Analysis

This report focused on reading several data sets, performing exploratory data analysis and saving a tidy data set to use for modeling.

List of [https://vincentarelbundock.github.io/Rdatasets/datasets.html](Data Sources)

We want to find something with at least 10 features, and 5000 observations
[https://archive.ics.uci.edu/ml/datasets.html?format=&task=cla&att=&area=&numAtt=10to100&numIns=&type=&sort=instDown&view=table]

* [https://archive.ics.uci.edu/ml/datasets/Buzz+in+social+media+](Buzz in Social Media)

* [Sensorless drive](https://archive.ics.uci.edu/ml/datasets/Dataset+for+Sensorless+Drive+Diagnosis) - 49 attributes, 11 classes, 58000 records

### Credit Payment Dataset
This comes from [card payment](https://archive.ics.uci.edu/ml/datasets/default+of+credit+card+clients) binary classifier with 23 predictors

```{r loaddata, echo=FALSE}

ccDataURL <- "https://archive.ics.uci.edu/ml/machine-learning-databases/00350/default%20of%20credit%20card%20clients.xls"
ccDataFile <- "./data/ccDataFile.xls"
dir.create("data", showWarnings=FALSE)
if (!file.exists(ccDataFile)) {
    download.file(url=ccDataURL, destfile=ccDataFile, mode="wb")
}
ccData <- read.xlsx2(file=ccDataFile, sheetIndex=1, startRow=2, stringsAsFactors=FALSE)


```

The basic approach will be to ensure no zero elements
NAndx <- sapply(data, function(x) mean(is.na(x))) > .9
nzv <- nearZeroVar(data)

```{r pulldata, echo=FALSE, warning=FALSE}

fCols <- c("SEX", "EDUCATION", "MARRIAGE", "default.payment.next.month")
nCols <- c("LIMIT_BAL", "AGE", "PAY_0", "PAY_2", "PAY_3", "PAY_4", "PAY_5", "PAY_6",
           "BILL_AMT1", "BILL_AMT2", "BILL_AMT3", "BILL_AMT4", "BILL_AMT5", "BILL_AMT6", 
           "PAY_AMT1", "PAY_AMT2", "PAY_AMT3", "PAY_AMT4", "PAY_AMT5", "PAY_AMT6" )
ccData <- select(ccData, -(ID)) %>% 
    mutate_at(fCols, funs(factor(.))) %>%
    mutate_at(nCols, funs(as.numeric(.)))

# go ahead and write it out for future notebooks
save(ccData, file="data/ccData.RData")
summary(ccData)
```

Some tables
```{r ccTables}
hist(ccData$LIMIT_BAL)

#hist(ccData$SEX)
#hist(ccData$MARRIAGE)
hist(ccData$AGE)

hist(ccData$PAY_0)
```


## Plots

Here we can see that if the bill amounts are much higher, more likely to default
```{r plot}
#ggplot(ccData, aes(PAY_AMT1, fill=MARRIAGE)) +
#    geom_density(alpha=0.3)
qplot(PAY_AMT1,BILL_AMT1,data=ccData, col=default.payment.next.month)
qplot(PAY_AMT2,BILL_AMT2,data=ccData, col=default.payment.next.month)
qplot(PAY_AMT3,BILL_AMT3,data=ccData, col=default.payment.next.month)

```

Lets look at the table of SEX, Marriage, education to the *default.payment.next.month* indicator.



## Save Formatted Data
Finally Save the datasets
```{r writeData}
save(ccData, file="ccData.RData")

```