---
title: "BasicDataAnalysis"
author: "Matt Cliff"
date: "January 29, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(tidyverse)
```

## Basic Data Analysis

This report focused on reading several data sets, performing exploratory data analysis and saving a tidy data set to use for modeling.


We want to find something with at least 10 features, and 5000 observations
[https://archive.ics.uci.edu/ml/datasets.html?format=&task=cla&att=&area=&numAtt=10to100&numIns=&type=&sort=instDown&view=table]

* [https://archive.ics.uci.edu/ml/datasets/Buzz+in+social+media+](Buzz in Social Media)

* [https://archive.ics.uci.edu/ml/datasets/Dataset+for+Sensorless+Drive+Diagnosis](Sensorless drive) - 49 attributs, 11 classes, 58000 records

* [https://archive.ics.uci.edu/ml/datasets/default+of+credit+card+clients](card payment) binary classifier with 23 predictors

```{r loaddata}
training <- data.frame(a=c(1,2,3),b=c(1,2,3))


```

The basic approach will be to ensure no zero elements
NAndx <- sapply(data, function(x) mean(is.na(x))) > .9
nzv <- nearZeroVar(data)

```{r pulldata}
ccDataURL <- "https://archive.ics.uci.edu/ml/machine-learning-databases/00350/default%20of%20credit%20card%20clients.xls"
ccDataFile <- "ccDataFile.xls"

download.file(url=ccDataURL, destfile=ccDataFile)
ccData <- readxl
```

## Including Plots

You can also embed plots, for example:


Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
