---
title: "My Neighborhood"
author: "Matt Cliff"
date: "April 3, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = NA)
require(dplyr)
require(ggplot2)
require(leaflet)
require(tibble)
```

## Neighborhood Charts

This notebook is in support of an application to be able to overlay certain geospatial data in and around my neighborhood in south Denver.

Based on [Leaflet](https://rstudio.github.io/leaflet/)

```{r intro}
df <- data.frame(lng=c(-104.985), lat=c(39.765))

m <- leaflet(df) %>% 
    addTiles() %>%
    addMarkers(lat=39.6445, lng=-104.9065, popup="Home")

#%>%
#    addProviderTiles(providers$Esri.WorldImagery)

m
```
perfect, now we have a large list of providers we can choose from
```{r providerList}
names(providers)
```


Let's add some nearby trail access points
```{r mkdataset}
trail_str = 
    "lat lng name trail
39.6568 -104.9034 'goldsmith' green
39.6426 -104.9408 'CO Highline' red
39.6529 -104.9466 'Hampden Highline' red
39.6596 -104.9403 'Eisenhower Pool' red
39.6594 -104.9317 'Dahlia Highline' red
39.6675 -104.9211 'Yale-1 Highline' red
39.6675 -104.9115 'Yale-2 Highline' red
39.6675 -104.9049 'Yale-3 Highline' red
39.6683 -104.8921 'Yale-4 Highline' red
39.6744 -104.9034 'Quebec Highline' red
39.6641 -104.8845 'Yosemite Highline' red
39.6386  -104.9361  'Highline Qunicy/Cherry Hills' red
    39.6350  -104.9322  'Highline Dahlia/Cherry Hills' red
     39.6599  -104.9035 'Highline/Goldsmith' green
     39.6347  -104.8930 'I225/Goldsmith' green
     39.6316  -104.8910 'Wallace Park' green
     39.6426  -104.8956 'Rosamund' green
     39.6239  -104.8890 'Belleview' green
     39.6440  -104.8803 'Village Green Trail' blue
      39.6615 -104.8778 'Hampden Heights' blue
      39.6277 -104.8758 'Campus' blue
      39.6601 -104.8649 'Cherry Creek-1' blue
      39.6322 -104.8755 'Cherry Creek-2' blue"



trails <- read.table(text=trail_str,
                     header=TRUE,
                     stringsAsFactors = FALSE) 
trails

m2 <- leaflet(trails) %>% 
    addTiles() %>%
    addAwesomeMarkers(~lng, ~lat, label=~name, popup=~name,  icon=awesomeIcons(icon="bicycle", library="fa", markerColor=~trail)) %>%
    addProviderTiles(providers$Esri.WorldImagery)

m2



t_str = 
    "lat lng name
     39.6445  -104.9065  'Home'"



homes <- read.table(text=t_str,
                     header=TRUE,
                     stringsAsFactors = FALSE)

m3 <- m2 %>% 
    addMarkers(lng = homes$lng, lat = homes$lat)

m3
```

Ok that is looking good

## Load Denver data

3 data sets

* street lights
* [311 calls](https://www.denvergov.org/opendata/dataset/city-and-county-of-denver-311-service-requests-2007-to-current)
* [foodstore locations](https://www.denvergov.org/opendata/dataset/city-and-county-of-denver-food-stores)
* [addresses lat/lng to denver address](https://www.denvergov.org/opendata/dataset/city-and-county-of-denver-addresses)
