---
title: "My Neighborhood"
author: "Matt Cliff"
date: "April 3, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = NA)
require(dplyr)
require(ggplot2)
require(leaflet)
require(tibble)
require(tidyr)
require(htmltools)

```

## Neighborhood Charts

This notebook is in support of an application to be able to overlay certain geospatial data in and around my neighborhood in south Denver.

Based on [Leaflet](https://rstudio.github.io/leaflet/)

```{r intro}
df <- data.frame(lng=c(-104.985), lat=c(39.765))

m <- leaflet(df) %>% 
    addTiles() %>%
    addMarkers(lat=39.6445, lng=-104.9065, popup="Home")

#%>%
#    addProviderTiles(providers$Esri.WorldImagery)

m
```
perfect, now we have a large list of providers we can choose from
```{r providerList}
names(providers)
```


Let's add some nearby trail access points
```{r mkdataset}
trail_str = 
    "lat lng name trail
39.6568 -104.9034 'goldsmith' green
39.6426 -104.9408 'CO Highline' red
39.6529 -104.9466 'Hampden Highline' red
39.6596 -104.9403 'Eisenhower Pool' red
39.6594 -104.9317 'Dahlia Highline' red
39.6675 -104.9211 'Yale-1 Highline' red
39.6675 -104.9115 'Yale-2 Highline' red
39.6675 -104.9049 'Yale-3 Highline' red
39.6683 -104.8921 'Yale-4 Highline' red
39.6744 -104.9034 'Quebec Highline' red
39.6641 -104.8845 'Yosemite Highline' red
39.6386  -104.9361  'Highline Qunicy/Cherry Hills' red
    39.6350  -104.9322  'Highline Dahlia/Cherry Hills' red
     39.6599  -104.9035 'Highline/Goldsmith' green
     39.6347  -104.8930 'I225/Goldsmith' green
     39.6316  -104.8910 'Wallace Park' green
     39.6426  -104.8956 'Rosamund' green
     39.6239  -104.8890 'Belleview' green
     39.6440  -104.8803 'Village Green Trail' blue
      39.6615 -104.8778 'Hampden Heights' blue
      39.6277 -104.8758 'Campus' blue
      39.6601 -104.8649 'Cherry Creek-1' blue
      39.6322 -104.8755 'Cherry Creek-2' blue"



trails <- read.table(text=trail_str,
                     header=TRUE,
                     stringsAsFactors = FALSE) 
trails

m2 <- leaflet(trails) %>% 
    addTiles() %>%
    addAwesomeMarkers(~lng, ~lat, label=~name, popup=~name,  icon=awesomeIcons(icon="bicycle", library="fa", markerColor=~trail)) %>%
    addProviderTiles(providers$Esri.WorldImagery)

m2



t_str = 
    "lat lng name
     39.6445  -104.9065  'Home'"



homes <- read.table(text=t_str,
                     header=TRUE,
                     stringsAsFactors = FALSE)

m3 <- m2 %>% 
    addMarkers(lng = homes$lng, lat = homes$lat)

m3
```

Ok that is looking good

## Load Denver data

3 data sets

* [street lights](https://www.denvergov.org/opendata/dataset/city-and-county-of-denver-street-light-poles)
* [311 calls](https://www.denvergov.org/opendata/dataset/city-and-county-of-denver-311-service-requests-2007-to-current)
* [foodstore locations](https://www.denvergov.org/opendata/dataset/city-and-county-of-denver-food-stores)
* **reference** [addresses lat/lng to denver address](https://www.denvergov.org/opendata/dataset/city-and-county-of-denver-addresses)


### Street Lights

```{r loadLightData, warning=FALSE}
datafile <- "streetlight.csv"
fileURL <- "https://www.denvergov.org/media/gis/DataCatalog/street_light_poles/csv/street_light_poles.csv"
if(!file.exists(datafile)) {
    download.file(fileURL, destfile=datafile)
}
streetlights <- read.csv(datafile)
names(streetlights)

#filter(lat < 39.763 & lat > 39.755) %>%
#    filter(lng > -104.989 & lng < -104.981) %>%


sldf <- subset(streetlights, 
               select=c(ESRIGNSS_LATITUDE, ESRIGNSS_LONGITUDE, ROAD_TYPE, POLE_TYPE, POLE_HT, POWER_SRC, DIMMER, ESRIGNSS_ALTITUDE, OWNER, LOCATION, INSTALL_YEAR, POLE_COND, LUMIN_WATT)) %>%
    rename(lat=ESRIGNSS_LATITUDE, lng=ESRIGNSS_LONGITUDE) %>%
    filter(lat > 39.75 & lat < 39.77) %>%
    filter(lng > -105.00 & lng < -104.96) %>%
    extract(col="POLE_HT", into="ht", "([0-9]+)" ) %>%
    extract(col="LUMIN_WATT", into="watts", "([0-9]+)" ) %>%
    mutate(ht = as.numeric(ht)) %>%
    mutate(watts = as.numeric(watts)) %>%
    mutate(ps = ifelse(POWER_SRC == "Underground", "Underground", "Above Ground"))


summary(sldf$ps)
```


The color represents if the powersource is underground or not, the size is propotional to the Lumin Wattage of the light.
```{r cityLightOverview, warning=FALSE}

pal <- colorFactor("Accent", domain=sldf$ps)

label0 <- paste("Pole Type:", sldf$POLE_TYPE)
label1 <- paste("Wattage:", sldf$watts)
labels <- paste(label0,paste("\n", label1))

leaflet(sldf) %>%
    addTiles() %>%
    addCircleMarkers(~lng, ~lat,
               weight = 1,
               radius= sldf$watts /20, 
               fillOpacity=0.3, 
               label = ~htmlEscape(labels),
               color = ~pal(ps),
               clusterOptions = markerClusterOptions()
               ) %>%
    addLegend("bottomright", pal = pal, values=sldf$ps, title="Power Supply") %>%
    setView(lat=39.765, lng=-104.985, zoom = 15) %>%
    addProviderTiles(providers$Esri.WorldImagery)

```

### Load the 311 data



```{r load311Data, warning=FALSE}
datafile <- "311service.csv"
fileURL <- "https://www.denvergov.org/media/gis/DataCatalog/311_service_requests_2007_to_current/csv/311_service_requests_2007_to_current.csv"

# 2016 only
fileURL2 <- "https://www.denvergov.org/media/gis/DataCatalog/311_service_data/csv/311_service_data_2016.csv"


if(!file.exists(datafile)) {
    download.file(fileURL, destfile=datafile)
}
raw_infocalls <- read.csv(datafile)
names(raw_infocalls)
dim(raw_infocalls)

infocalls <- raw_infocalls %>%
   rename(lat=Latitude, lng=Longitude) %>%
    filter(lat > 39.75 & lat < 39.77) %>%
    filter(lng > -105.00 & lng < -104.96)


```
ok... now in business

```{r three11Overview, warning=FALSE}
leaflet(infocalls) %>%
    addTiles() %>%
    addCircleMarkers(~lng, ~lat,
               weight = 1,
               fillOpacity=0.3, 
               clusterOptions = markerClusterOptions()
               ) %>%
    setView(lat=39.7, lng=-104.9, zoom = 10) %>%
    addProviderTiles(providers$Esri.WorldImagery)

```





### grocery data


```{r loadFoodData, warning=FALSE}
datafile <- "food.csv"
fileURL <- "https://www.denvergov.org/media/gis/DataCatalog/food_stores/csv/food_stores.csv"


if(!file.exists(datafile)) {
    download.file(fileURL, destfile=datafile)
}
raw_food <- read.csv(datafile)
names(raw_food)
summary(raw_food)

food <- raw_food %>%
   rename(lat=POINT_Y, lng=POINT_X, label=STORE_NAME)
table(food$STORE_TYPE)
```

Basic Food plot
```{r foodOverview, warning=FALSE}

pal <- colorFactor("Accent", domain=food$STORE_TYPE)


leaflet(food) %>%
    addTiles() %>%
    addCircleMarkers(~lng, ~lat, 
               weight = 1,
               fillOpacity = 0.3, 
               color = ~pal(STORE_TYPE)
               ) %>%
        setView(lat=39.7, lng=-104.9, zoom = 10) %>%
addProviderTiles(providers$Esri.WorldImagery)

```

## Write Output

Write out the datasets for usage in the shiny app

```{r saveDatasets}
lights <- sldf
save(trails, file="trails.RData")
save(lights, file="lights.RData")
save(infocalls, file="infocalls.RData")
save(food, file="food.RData")